---
title: "Advanced Statistical Computing: Endterm Project 2020"
output: html_document
---

### Name: Deniz Sen

### ULCN: 1486438

```{r label = preps, echo=FALSE}
rm(list=ls()) #clears your workspace
```
### Load data and Packages

```{r}
getwd()
library(copula)
data = read.csv("insurance.csv")
data = data[,-1]
x1 = data[,1]
x2 = data[,2]
```

### Illustrate the dependence in the data with a graph.
```{r}
# Graph Dependency 
plot(data[,1], data[,2], xlim = c(0, 30), ylim = c(0, 50), pch = 1,  cex = 1,
     col = "darkblue",
     main = "Scatterpot: PLI vs WC" ,xlab = "PLI", ylab = "WC")
abline(lm(data[,2] ~ data[,1]), col='red', lwd = 2)
legend("topright", "Linear Relation", 
       col = "red", lty = 1, lwd = 2, cex = 0.70)
cor(data[,1], data[,2], method ="pearson") 

# Graph Price as a function of threshold t
t = t = seq(from = 100, to = 200, by = 10) #threshold t
P = function(t) {
  40000 * exp(-t / 7) 
}
plot(t, P(t), type = "l", lwd= 2,  main = "Price as a function of threshold t",
     xlab = "Threshold t", ylab = "Price", col ="darkblue")
legend("topright", "P(t)", 
       col = "darkblue", lty = 1, lwd = 2, cex = 0.70)
```

### Maximum Likelihood estimation of f_x_j , j = 1,2

```{r}
theta = rbind( c(mean(x1), sd(x1)), c(mean(x2), sd(x2)))
theta
log_l = function(theta, data){
  mu = theta[1]
  sigma = theta[2]
  logl = suppressWarnings(-sum(log(dlnorm(data, meanlog = mu, sdlog = sigma))))
  return(logl)
}
res_optim = matrix(0, nrow = nrow(theta), ncol = 2)
for(i in 1: nrow(theta)) {
  res_optim[i,] = optim(par = theta[i,], fn = log_l, data = data[,i], method="BFGS")$par
}
colnames(res_optim) = c("mu", "sigma")
res_optim
```

### Function that fits the parameter of a Joe copula model.
```{r}
copula_mod = function (theta, data) {
  res_optim_temp = matrix(0, nrow = 1, ncol = 2)
  Fx = matrix(0, nrow = nrow(data), ncol = 2)
  for(i in 1: nrow(theta)) {
    res_optim_temp = optim(par = theta[i,], fn = log_l, data = data[,i], method="BFGS")$par
    Fx[,i] = plnorm(data[,i], meanlog = res_optim_temp[1], sdlog = res_optim_temp[2])
  }
  U = Fx #pseudo observations
  return(U)
}

# generate pseudo observations
U = copula_mod(theta, data)

# log likelihood of copula density
copula_log_l = function(theta, data) {
  -sum(log(dCopula(data, joeCopula(theta))))
}
res_optim_copula = t(matrix(optimize(copula_log_l, data = U, lower = 1e-6, upper = 5)))
colnames(res_optim_copula) = c("copula_theta", "max_log_l_hood") # standard error corresponding max LL

# Compare values using fitCopula() 
fit_copula = fitCopula(joeCopula(), data = U)
summary(fit_copula)
res_optim_copula  # results are the same
```


### Function that estimates all model parameters
```{r}
copula_fit = function(theta, data){
  #i. & ii. Compute estimates (mu_j, sigma_j), j =1,2 by maximum likelihood.
  res_mle = matrix(0, nrow = nrow(theta), ncol = 2)
  for(i in 1: nrow(theta)) {
    res_mle[i,] = optim(par = theta[i,], fn = log_l, data = data[,i], method="BFGS")$par
  }
  colnames(res_mle) = c("mu", "sigma")
  # iii. Set U_j = F_mu_j,sigma_j (X_j),j=1,2 and estimate the parameters of 
  # a Joe copula model for (U_1,U_2)
  U = copula_mod(theta, data) # helper function "copula_mod" defined above
  
  res_theta_copula = t(matrix(unlist(optimize(copula_log_l, data = U, lower = 1e-6, upper = 5))))
  colnames(res_theta_copula) = c("sd_err", "MLL") # standard error corresponding max
  all_thetas = as.matrix(c(res_mle, res_theta_copula)) 
  row.names(all_thetas) = c("mle_mu1","mle_mu2", "mle_sigma1", "mle_sigma2", 
                       "copula_theta", "copula_ml")
  colnames(all_thetas) = "estimates"
  return(all_thetas)
}
fit = copula_fit(theta, data)
fit
theta
```

### Function that simulates from the joint model for (X_1,X_2)
```{r}
sim_join_mod_obs = function(mu1, sigma1, mu2, sigma2, theta, n) {
  sim = rCopula(n, joeCopula(theta))
  # transform uniform to log-normal
  x1 = qlnorm(sim[, 1], meanlog = mu1, sdlog = sigma1)
  x2 = qlnorm(sim[, 2], meanlog = mu2, sdlog = sigma2)
  return(cbind(x1, x2))
}
n = dim(data)[1]
sim_join_res = sim_join_mod_obs(mu1 = fit[1,], mu2= fit[2,], 
             sigma1=fit[3,], sigma2=fit[4,], 
             theta = fit[5,], n = n)
```

### Compare the simulated data with the observed data. 
```{r}

plot(x = data[,1], y = data[,2], pch = 1, cex = 1, col = "blue",
     main = "Scatterplot: Observed vs. Simulated data ", xlab = "PLI", 
     ylab = "WC")
points(sim_join_res[,1], sim_join_res[,2], pch = 3, cex=.7, col = "red")
legend("bottomright", c("Observed", "Simulated"),
        col = c("blue", "red"), pch = c(1,3), cex = 0.70)

hist(data[,1], breaks = 50, prob = TRUE,  
     main = "Histogram Observed PLI vs. Simulated PLI Variable",
     xlab = "PLI")
lines(density(data[,1]), type = "l", lwd = 3, lty = 2, col = "blue")
lines(density(sim_join_res[,1]), type = "l", lwd = 3, lty = 3, col = "red")
legend("topright", c("Observed", "Simulated"), 
       col = c("blue", "red"), lty = c(2,3), 
       lwd=3, cex = 0.70)

hist(data[,2], breaks = 50, prob = TRUE,
     main = "Histogram Observed WC vs. Simulated WC Variable",
     xlab = "WC")
lines(density(data[,2]), type = "l", lwd = 3, lty = 2, col = "blue")
lines(density(sim_join_res[,2]), type = "l",lwd = 3, lty = 3, col = "red")
legend("topright", c("Observed", "Simulated"), 
       col = c("blue", "red"), lty = c(2,3), 
       lwd = 3, cex = 0.70)
```

### Ilustrate how properties of the data change when paraemters are de/increased
```{r}
sim_join_res = sim_join_mod_obs(mu1 = fit[1,], mu2= fit[2,], 
             sigma1=fit[3,], sigma2=fit[4,], 
             theta = fit[5,]+3, n = n)

plot(x = data[,1], y = data[,2], pch = 1, cex = 1, col = "blue",
     main = "Scatterplot: Observed vs. Simulated data ", xlab = "PLI", 
     ylab = "WC")
points(sim_join_res[,1], sim_join_res[,2], pch = 3, cex=.7, col = "red")
legend("bottomright", c("Observed", "Simulated"),
        col = c("blue", "red"), pch = c(1,3), cex = 0.70)

hist(data[,1], breaks = 50, prob = TRUE,  
     main = "Histogram Observed PLI vs. Simulated PLI Variable",
     xlab = "PLI")
lines(density(data[,1]), type = "l", lwd = 3, lty = 2, col = "blue")
lines(density(sim_join_res[,1]), type = "l", lwd = 3, lty = 3, col = "red")
legend("topright", c("Observed", "Simulated"), 
       col = c("blue", "red"), lty = c(2,3), 
       lwd=3, cex = 0.70)

hist(data[,2], breaks = 50, prob = TRUE,
     main = "Histogram Observed WC vs. Simulated WC Variable",
     xlab = "WC")
lines(density(data[,2]), type = "l", lwd = 3, lty = 2, col = "blue")
lines(density(sim_join_res[,2]), type = "l",lwd = 3, lty = 3, col = "red")
legend("topright", c("Observed", "Simulated"), 
       col = c("blue", "red"), lty = c(2,3), 
       lwd = 3, cex = 0.70)
```

```{r}
# Illustrate how the properties of the data change when you increase/decrease a 
# parameter.
sim_join_res = sim_join_mod_obs(mu1 = fit[1,], mu2= fit[2,], 
             sigma1=fit[3,], sigma2=fit[4,], 
             theta = 1, n = n)

plot(x = data[,1], y = data[,2], pch = 1, cex = 1, col = "blue",
     main = "Scatterplot: Observed vs. Simulated data ", xlab = "PLI", 
     ylab = "WC")
points(sim_join_res[,1], sim_join_res[,2], pch = 3, cex=.7, col = "red")
legend("bottomright", c("Observed", "Simulated"),
        col = c("blue", "red"), pch = c(1,3), cex = 0.70)

hist(data[,1], breaks = 50, prob = TRUE,  
     main = "Histogram Observed PLI vs. Simulated PLI Variable",
     xlab = "PLI")
lines(density(data[,1]), type = "l", lwd = 3, lty = 2, col = "blue")
lines(density(sim_join_res[,1]), type = "l", lwd = 3, lty = 3, col = "red")
legend("topright", c("Observed", "Simulated"), 
       col = c("blue", "red"), lty = c(2,3), 
       lwd=3, cex = 0.70)

hist(data[,2], breaks = 50, prob = TRUE,
     main = "Histogram Observed WC vs. Simulated WC Variable",
     xlab = "WC")
lines(density(data[,2]), type = "l", lwd = 3, lty = 2, col = "blue")
lines(density(sim_join_res[,2]), type = "l",lwd = 3, lty = 3, col = "red")
legend("topright", c("Observed", "Simulated"), 
       col = c("blue", "red"), lty = c(2,3), 
       lwd = 3, cex = 0.70)
```

```{r}
sim_join_res = sim_join_mod_obs(mu1 = fit[1,]+.5, mu2= fit[2,]+.5, 
             sigma1=fit[3,], sigma2=fit[4,], 
             theta = fit[5,], n = n)

plot(x = data[,1], y = data[,2], pch = 1, cex = 1, col = "blue",
     main = "Scatterplot: Observed vs. Simulated data ", xlab = "PLI", 
     ylab = "WC")
points(sim_join_res[,1], sim_join_res[,2], pch = 3, cex=.7, col = "red")
legend("bottomright", c("Observed", "Simulated"),
        col = c("blue", "red"), pch = c(1,3), cex = 0.70)

hist(data[,1], breaks = 50, prob = TRUE,  
     main = "Histogram Observed PLI vs. Simulated PLI Variable",
     xlab = "PLI")
lines(density(data[,1]), type = "l", lwd = 3, lty = 2, col = "blue")
lines(density(sim_join_res[,1]), type = "l", lwd = 3, lty = 3, col = "red")
legend("topright", c("Observed", "Simulated"), 
       col = c("blue", "red"), lty = c(2,3), 
       lwd=3, cex = 0.70)

hist(data[,2], breaks = 50, prob = TRUE,
     main = "Histogram Observed WC vs. Simulated WC Variable",
     xlab = "WC")
lines(density(data[,2]), type = "l", lwd = 3, lty = 2, col = "blue")
lines(density(sim_join_res[,2]), type = "l",lwd = 3, lty = 3, col = "red")
legend("topright", c("Observed", "Simulated"), 
       col = c("blue", "red"), lty = c(2,3), 
       lwd = 3, cex = 0.70)
```

```{r}
sim_join_res = sim_join_mod_obs(mu1 = fit[1,], mu2= fit[2,], 
             sigma1=fit[3,]*3, sigma2=fit[4,]*3, 
             theta = 1, n = n)

plot(x = data[,1], y = data[,2], pch = 1, cex = 1, col = "blue",
     main = "Scatterplot: Observed vs. Simulated data ", xlab = "PLI", 
     ylab = "WC")
points(sim_join_res[,1], sim_join_res[,2], pch = 3, cex=.7, col = "red")
legend("bottomright", c("Observed", "Simulated"),
        col = c("blue", "red"), pch = c(1,3), cex = 0.70)

hist(data[,1], breaks = 50, prob = TRUE,  
     main = "Histogram Observed PLI vs. Simulated PLI Variable",
     xlab = "PLI")
lines(density(data[,1]), type = "l", lwd = 3, lty = 2, col = "blue")
lines(density(sim_join_res[,1]), type = "l", lwd = 3, lty = 3, col = "red")
legend("topright", c("Observed", "Simulated"), 
       col = c("blue", "red"), lty = c(2,3), 
       lwd=3, cex = 0.70)

hist(data[,2], breaks = 50, prob = TRUE,
     main = "Histogram Observed WC vs. Simulated WC Variable",
     xlab = "WC")
lines(density(data[,2]), type = "l", lwd = 3, lty = 2, col = "blue")
lines(density(sim_join_res[,2]), type = "l",lwd = 3, lty = 3, col = "red")
legend("topright", c("Observed", "Simulated"), 
       col = c("blue", "red"), lty = c(2,3), 
       lwd = 3, cex = 0.70)
```


### Simulation study to understand inner workins of method
### Fix mu1 = 1, sigma1 = 2, mu2 = 3, sigma2 = 0.5, theta = 2. 
### For r = 1,...,100: repetitions on n = 200,500,1000 simulated observations
```{r}
n = as.vector(c(200, 500, 1000))
R = 100 # 100 repetitions
theta8 = rbind( c(1, 2), c(3, 0.5))
rep_res = list()
rep_fit = matrix(0, nrow = R, ncol = 6)
colnames(rep_fit) = c("mle_mu1","mle_mu2", "mle_sigma1", "mle_sigma2", 
                          "copula_sde", "max_log_l_hood")
rmse_m = matrix(0, nrow = length(n), ncol = 6)
colnames(rmse_m) = c("mle_mu1","mle_mu2", "mle_sigma1", "mle_sigma2", 
                      "copula_sde", "max_log_l_hood")
row.names(rmse_m) = c("n = 200", "n = 500", "n = 1000")
time = matrix(0, nrow = R, ncol = length(n))

rmse = function(x) {
  sd(x) / sqrt(n[i])
}

for (i in 1:length(n)){
  for (r in 1:R){
    U_obs = sim_join_mod_obs(mu1 = 1, mu2 = 3, 
                             sigma1 = 2, sigma2 = 0.5, 
                             theta = 2, n = n[i])
    time[r,i] <- system.time(rep_fit[r,] <- copula_fit(theta = theta8,  data = U_obs))[3]
    rep_res[i] = list(rep_fit)
    names(rep_res)[i] = paste0("n", n[i])
  }
  rmse_m[i, ] = apply(rep_fit, 2, FUN = rmse)
}

plot(n, rmse_m[,1], ylim = c(0, max(rmse_m[,1:5])), pch = 19, 
     col = "dark slate blue",  xlab ="Number of observations n",
     ylab="RMSE", main = "RMSE as a function of sample size n",
     cex.main = 1)
lines(n,rmse_m[,1], type = "b", col = "darkblue")
points(n,rmse_m[,2], pch = 19, col = "medium slate blue")
lines(n,rmse_m[,2], type = "b", col = "medium slate blue")
points(n, rmse_m[,3], pch = 19, col = "medium sea green")
lines(n,rmse_m[,3], type = "b", col = "medium sea green")
points(n, rmse_m[,4], pch = 19, col = "light sea green")
lines(n,rmse_m[,4], type = "b", col = "light sea green")
points(n, rmse_m[,5], pch = 19, col = "red")
lines(n,rmse_m[,5], type = "b", col = "red")
legend("topright", c(expression(mu[1]), expression(mu[2]), 
                     expression(sigma[1]),expression(sigma[2]),
                     expression(theta)),
       col = c("darkblue", "medium slate blue",
               "medium sea green","light sea green",
               "red"), lty = c(1,1), 
       lwd = 3, cex = 0.70)

# plot mean computation time
mean_time = colMeans(time)
mean_time
plot(x = n, y = mean_time, col = "darkblue", 
     main = "Mean computation time as a function of n", 
     xlab = "Number of Observations n", ylab = "Mean computation time")
lines(x = n, mean_time,col = "darkblue", lwd = 2)

```

### Fit all paremeter models and compute expected payout using MCMC simulation
```{r}
# fit all model parameters to the observed data
fit = copula_fit(theta, data)
n = 10^5
t = seq(from = 100, to = 200, by = 10) #threshold t
V = numeric(length(t))
# price function
P = function(t) {
  40000 * exp(-t / 7) 
}
# function for expected reinsurance payout
f = function(x1, x2, t) { 
  (x1 + x2 > t) * (x1 + x2)
}

for (i in 1:length(t)){
  U_obs = sim_join_mod_obs(mu1 = fit[1,], mu2 = fit[2,], 
                                   sigma1 = fit[3,], sigma2 = fit[4,], 
                                   theta = fit[5,], n = n)
  # compute V(t)
  V[i] = mean(f(x1 = U_obs[,1], x2 = U_obs[,2], t = t[i]))
}

plot(x = t, y = V, pch = 1, cex= 1, col = "red",
     main = "Expected Reinsurance Payout V(t) and Price P(t)", xlab = "Treshold t", 
     ylab = "Insurance Payout")
lines(t, V, type = "b", lwd = 2, lty = 3, col = "red")
lines(t, P(t), type = "b", lwd = 2, lty = 2, col = "darkblue")
legend("topright", c("V(t)", "P(t)"), 
       col = c("red", "darkblue"), lty = c(3,2), 
       lwd = 2, cex = 0.70)
```

### Compute expected reinsurance payout using importance sampling
```{r}
# V(t) are quite noisy. Compute the values again using Importance Sampling
joint = function(x1, x2, mu1, sigma1, mu2, sigma2, theta) {
  fx1 = dlnorm(x1, meanlog = mu1, sdlog = sigma1)
  fx2 = dlnorm(x2, meanlog = mu2, sdlog = sigma2)
  Fx1 = plnorm(x1, meanlog = mu1, sdlog = sigma1)
  Fx2 = plnorm(x2,  meanlog = mu2, sdlog = sigma2)
  U = cbind(Fx1, Fx2)
  copula_res = dCopula(U, joeCopula(theta))
  dens = fx1 * fx2 * copula_res # joint density = prod of marginals and copula
  return(dens)
}

theta = rbind( c(mean(x1), sd(x1)), c(mean(x2), sd(x2)))
fit = copula_fit(theta, data)
mu1 = fit[1,]; mu2 = fit[2,]; sigma1 = fit[3,]; sigma2 = fit[4,]
copula_theta = fit[5,]

g = function(x1, x2, t, fx_y, fy_y) {
  (x1 + x2) * (x1 + x2 > t) * fx_y / fy_y
}

# price function 
P = function(t) {
  40000 * exp(-t / 7) 
}

n = 10^5
V_res = numeric(length(t))
for (i in 1:length(t)) {
  U_obs = sim_join_mod_obs(mu1 = 2+mu1, mu2 = 2+mu2, 
                       sigma1 = sigma1, sigma2 = sigma2, 
                       theta = copula_theta, n = n)
  fx_y = joint(U_obs[, 1], U_obs[, 2], mu1, sigma1, mu2, sigma2, copula_theta)
  fy_y = joint(U_obs[, 1], U_obs[, 2], 2 + mu1, sigma1, 2+mu2, sigma2, copula_theta)
  V_res[i] = mean(g(U_obs[, 1], U_obs[, 2], t[i], fx_y, fy_y))
}

# Regular scale
P = P(t) # price
plot(x = t, y = V_res, ylim = c(0, max(V_res)), pch = 1, cex= 1, 
     col = "red",
     main = "Importance Sampling: Expected Reinsurance Payout V(t) and Price P(t)", 
     cex.main= .9, xlab = "Treshold t", ylab = "Insurance Payout")
lines(t, V_res, type = "b", lwd = 2,lty = 3, col = "red")
lines(t, P, type = "b", lwd = 2,lty = 2, col = "darkblue")
legend("topright", c("V", "P"), 
       col = c("red", "darkblue"), lty = c(3,2), 
       lwd=2, cex = 0.70)

# log scale
logV = log(V_res)
logP = log(P)
plot(x = t, y = logV, ylim = c(min(logP), max(logP)), pch = 1, cex= 1,
     col = "red",
     main = "Importance Sampling: Log of Expected Reinsurance Payout V(t) and Price P(t)", 
     cex.main= .8, xlab = "Treshold t", ylab = "Insurance Payout")
lines(t, logV, type = "b", lwd = 2, lty = 3, col = "red")
lines(t, logP, type = "b", lwd = 2, lty = 2, col = "darkblue")
legend("topright", c("Log(V(t))", "Log(P(t))"), 
       col = c("red", "darkblue"), lty = c(3,2), 
       lwd = 2, cex = 0.70)
```

### Bootstrap method to compute 80% confidence intervals for expected payout
```{r}
# Use a bootstrap method to compute 80 % CI for V(t)
n = nrow(data)
B = 1000
V_temp = numeric(length(t))
V_boot = matrix(nrow = B, ncol = length(t))
theta = theta = rbind( c(mean(x1), sd(x1)), c(mean(x2), sd(x2)))

for (b in 1:B){
  i = sample(1:n, n, replace = T)
  X = data[i,1:2]
  modfit = copula_fit(theta, X)
  mu1 = modfit[1,]; mu2 = modfit[2,]; sigma1 = modfit[3,];  sigma2 = modfit[4,];
  copula_theta = modfit[5,]
  for (j in 1:length(t)){
  U_obs = sim_join_mod_obs(mu1 = mu1 + 2, mu2 =  mu2 + 2, 
                           sigma1 = sigma1, sigma2 = sigma2, 
                           theta = copula_theta, n = n)
  fx_y = joint(U_obs[, 1], U_obs[, 2], mu1, sigma1, mu2, sigma2, copula_theta)
  fy_y = joint(U_obs[, 1], U_obs[, 2], 2+mu1, sigma1, 2+mu2, sigma2, copula_theta)
  
  V_temp[j] = mean(g(U_obs[, 1], U_obs[, 2], t[j], fx_y, fy_y))
  }
  V_boot[b,] = V_temp
}
V_boot
low_bound = numeric(length(t))
up_bound = numeric(length(t))
for (i in 1:length(t)){
  low_bound[i] = sort(V_boot[,i])[B*0.10]
  up_bound[i] = sort(V_boot[,i])[B*0.90]
}
# min(low_bound)
# min(up_bound)
# min(V_res)
# min(P)
# max(P)
# max(V_res)
# max(up_bound)
# max(low_bound)

plot(x = t, y = V_res, ylim = c(min(P), max(up_bound)), pch = 1, cex= 1, 
     col = "red",
     main = "Bootstraped 80 % CI: Expected Reinsurance Payout V(t) and Price P(t)", 
     xlab = "Treshold t", ylab = "Insurance Payout", cex.main=.9)
lines(t, V_res, type = "b", lwd = 2, col = "red")
lines(t, P, type = "b", lwd = 2, , lty = 3, col = "darkblue")
segments(t, up_bound, t, low_bound, lwd = 2, col = "grey25")
arrows(t, up_bound, t, low_bound, code = 3, lwd = 2, col = "grey25", angle = 90, 
       length = 0.1)
legend("topright", c("V", "P", "80 % CI"), 
       col = c("red", "darkblue", "grey25"), lty = c(3,2,1), 
       lwd = 2, cex = 0.70)

# log scale 
loglow =  log(low_bound)
logup = log(up_bound)
# min(logup)
# min(loglow)
# min(logV)
# min(logP)
# max(logV)
# max(logP)
# max(logup)
# max(loglow)
plot(x = t, y = logV, ylim = c(min(logP), max(logup)), pch = 1, cex= 1, 
     col = "red",
     main = "Bootstraped 80 % CI: Expected Reinsurance Payout V(t) and Price P(t)", 
     xlab = "Treshold t", ylab = "Insurance Payout", cex.main=.9)
lines(t, logV, type = "b", lwd = 2, col = "red")
lines(t, logP, type = "b", lwd = 2, , lty = 3, col = "darkblue")
segments(t, loglow, t, logup, lwd = 2, col = "grey25")
arrows(t, loglow, t, logup, code = 3, lwd = 2, col = "grey25", angle = 90, 
       length = 0.1)
legend("topright", c("log(V)", "log(P)", "80 % CI"), 
       col = c("red", "darkblue", "grey25"), lty = c(3,2,1), 
       lwd = 2, cex = 0.70)
```

### Appendix for Exercise 8: Auxilliary Simulations
```{r}
# sigma 1 = sigma 2 = 0.5
n = as.vector(c(200, 500, 1000))
R = 100 # 100 repetitions
theta8 = rbind( c(1, 2), c(3, 0.5))
rep_res = list()
rep_fit = matrix(0, nrow = R, ncol = 6)
colnames(rep_fit) = c("mle_mu1","mle_mu2", "mle_sigma1", "mle_sigma2", 
                          "copula_sde", "max_log_l_hood")
rmse_m = matrix(0, nrow = length(n), ncol = 6)
colnames(rmse_m) = c("mle_mu1","mle_mu2", "mle_sigma1", "mle_sigma2", 
                      "copula_sde", "max_log_l_hood")
row.names(rmse_m) = c("n = 200", "n = 500", "n = 1000")
time = matrix(0, nrow = R, ncol = length(n))

rmse = function(x) {
  sd(x) / sqrt(n[i])
}

for (i in 1:length(n)){
  for (r in 1:R){
    U_obs = sim_join_mod_obs(mu1 = 1, mu2 = 3, 
                             sigma1 = 0.5, sigma2 = 0.5, 
                             theta = 2, n = n[i])
    time[r,i] <- system.time(rep_fit[r,] <- copula_fit(theta = theta8,  data = U_obs))[3]
    rep_res[i] = list(rep_fit)
    names(rep_res)[i] = paste0("n", n[i])
  }
  rmse_m[i, ] = apply(rep_fit, 2, FUN = rmse)
}

plot(n, rmse_m[,1], ylim = c(0, max(rmse_m[,1:5])), pch = 19, 
     col = "dark slate blue",  xlab ="Number of observations n",
     ylab="RMSE", main = "RMSE as a function of sample size n",
     cex.main = 1)
lines(n,rmse_m[,1], type = "b", col = "darkblue")
points(n,rmse_m[,2], pch = 19, col = "medium slate blue")
lines(n,rmse_m[,2], type = "b", col = "medium slate blue")
points(n, rmse_m[,3], pch = 19, col = "medium sea green")
lines(n,rmse_m[,3], type = "b", col = "medium sea green")
points(n, rmse_m[,4], pch = 19, col = "light sea green")
lines(n,rmse_m[,4], type = "b", col = "light sea green")
points(n, rmse_m[,5], pch = 19, col = "red")
lines(n,rmse_m[,5], type = "b", col = "red")
legend("topright", c(expression(mu[1]), expression(mu[2]), 
                     expression(sigma[1]),expression(sigma[2]),
                     expression(theta)),
       col = c("darkblue", "medium slate blue",
               "medium sea green","light sea green",
               "red"), lty = c(1,1), 
       lwd = 3, cex = 0.70)

# plot mean computation time
mean_time = colMeans(time)
mean_time
plot(x = n, y = mean_time, col = "darkblue", 
     main = "Mean computation time as a function of n", 
     xlab = "Number of Observations n", ylab = "Mean computation time")
lines(x = n, mean_time,col = "darkblue", lwd = 2)

```

```{r}
# theta  = 1.6
n = as.vector(c(200, 500, 1000))
R = 100 # 100 repetitions
theta8 = rbind( c(1, 2), c(3, 0.5))
rep_res = list()
rep_fit = matrix(0, nrow = R, ncol = 6)
colnames(rep_fit) = c("mle_mu1","mle_mu2", "mle_sigma1", "mle_sigma2", 
                          "copula_sde", "max_log_l_hood")
rmse_m = matrix(0, nrow = length(n), ncol = 6)
colnames(rmse_m) = c("mle_mu1","mle_mu2", "mle_sigma1", "mle_sigma2", 
                      "copula_sde", "max_log_l_hood")
row.names(rmse_m) = c("n = 200", "n = 500", "n = 1000")
time = matrix(0, nrow = R, ncol = length(n))

rmse = function(x) {
  sd(x) / sqrt(n[i])
}

for (i in 1:length(n)){
  for (r in 1:R){
    U_obs = sim_join_mod_obs(mu1 = 1, mu2 = 3, 
                             sigma1 = 2, sigma2 = 0.5, 
                             theta = 1.6, n = n[i])
    time[r,i] <- system.time(rep_fit[r,] <- copula_fit(theta = theta8,  data = U_obs))[3]
    rep_res[i] = list(rep_fit)
    names(rep_res)[i] = paste0("n", n[i])
  }
  rmse_m[i, ] = apply(rep_fit, 2, FUN = rmse)
}

plot(n, rmse_m[,1], ylim = c(0, max(rmse_m[,1:5])), pch = 19, 
     col = "dark slate blue",  xlab ="Number of observations n",
     ylab="RMSE", main = "RMSE as a function of sample size n",
     cex.main = 1)
lines(n,rmse_m[,1], type = "b", col = "darkblue")
points(n,rmse_m[,2], pch = 19, col = "medium slate blue")
lines(n,rmse_m[,2], type = "b", col = "medium slate blue")
points(n, rmse_m[,3], pch = 19, col = "medium sea green")
lines(n,rmse_m[,3], type = "b", col = "medium sea green")
points(n, rmse_m[,4], pch = 19, col = "light sea green")
lines(n,rmse_m[,4], type = "b", col = "light sea green")
points(n, rmse_m[,5], pch = 19, col = "red")
lines(n,rmse_m[,5], type = "b", col = "red")
legend("topright", c(expression(mu[1]), expression(mu[2]), 
                     expression(sigma[1]),expression(sigma[2]),
                     expression(theta)),
       col = c("darkblue", "medium slate blue",
               "medium sea green","light sea green",
               "red"), lty = c(1,1), 
       lwd = 3, cex = 0.70)

# plot mean computation time
mean_time = colMeans(time)
mean_time
plot(x = n, y = mean_time, col = "darkblue", 
     main = "Mean computation time as a function of n", 
     xlab = "Number of Observations n", ylab = "Mean computation time")
lines(x = n, mean_time,col = "darkblue", lwd = 2)
````

```{r}
# theta  = 5
n = as.vector(c(200, 500, 1000))
R = 100 # 100 repetitions
theta8 = rbind( c(1, 2), c(3, 0.5))
rep_res = list()
rep_fit = matrix(0, nrow = R, ncol = 6)
colnames(rep_fit) = c("mle_mu1","mle_mu2", "mle_sigma1", "mle_sigma2", 
                          "copula_sde", "max_log_l_hood")
rmse_m = matrix(0, nrow = length(n), ncol = 6)
colnames(rmse_m) = c("mle_mu1","mle_mu2", "mle_sigma1", "mle_sigma2", 
                      "copula_sde", "max_log_l_hood")
row.names(rmse_m) = c("n = 200", "n = 500", "n = 1000")
time = matrix(0, nrow = R, ncol = length(n))

rmse = function(x) {
  sd(x) / sqrt(n[i])
}

for (i in 1:length(n)){
  for (r in 1:R){
    U_obs = sim_join_mod_obs(mu1 = 1, mu2 = 3, 
                             sigma1 = 2, sigma2 = 0.5, 
                             theta = 5, n = n[i])
    time[r,i] <- system.time(rep_fit[r,] <- copula_fit(theta = theta8,  data = U_obs))[3]
    rep_res[i] = list(rep_fit)
    names(rep_res)[i] = paste0("n", n[i])
  }
  rmse_m[i, ] = apply(rep_fit, 2, FUN = rmse)
}

plot(n, rmse_m[,1], ylim = c(0, max(rmse_m[,1:5])), pch = 19, 
     col = "dark slate blue",  xlab ="Number of observations n",
     ylab="RMSE", main = "RMSE as a function of sample size n",
     cex.main = 1)
lines(n,rmse_m[,1], type = "b", col = "darkblue")
points(n,rmse_m[,2], pch = 19, col = "medium slate blue")
lines(n,rmse_m[,2], type = "b", col = "medium slate blue")
points(n, rmse_m[,3], pch = 19, col = "medium sea green")
lines(n,rmse_m[,3], type = "b", col = "medium sea green")
points(n, rmse_m[,4], pch = 19, col = "light sea green")
lines(n,rmse_m[,4], type = "b", col = "light sea green")
points(n, rmse_m[,5], pch = 19, col = "red")
lines(n,rmse_m[,5], type = "b", col = "red")
legend("topright", c(expression(mu[1]), expression(mu[2]), 
                     expression(sigma[1]),expression(sigma[2]),
                     expression(theta)),
       col = c("darkblue", "medium slate blue",
               "medium sea green","light sea green",
               "red"), lty = c(1,1), 
       lwd = 3, cex = 0.70)

# plot mean computation time
mean_time = colMeans(time)
mean_time
plot(x = n, y = mean_time, col = "darkblue", 
     main = "Mean computation time as a function of n", 
     xlab = "Number of Observations n", ylab = "Mean computation time")
lines(x = n, mean_time,col = "darkblue", lwd = 2)
````
